Synthesizing Quotient Lenses

### 2017-11-02 papers/QuotientLensSynth/main.tex
finished a pass through section 4

@@

\begin{figure}[t]
  \centering
\saz{Fix the projection case}
\[
    \begin{array}{l@{\quad}l@{\quad}l} 
      w \; \equiv_R \; w' &\iff& w = w' \\
      w \; \equiv_{R \mapsto s} \; w' \text{ for all }w, w'\\
      w \; \equiv_{\squash{R}{R'}{f}} \; w' &\iff& f(w) = w'
      \text{ or } w = w' \\
      w \; \equiv_{\normalize{R}{R'}{f}} \; w' &\iff&
      f(w)=f(w') \text{ or }w = w'\\
      w \; \equiv_{q_2 \circ q_1} \; w' &\iff& \exists k, k' \in
  \mathcal{L}(K(q_2)) \text{ such that } w \; \equiv_{q_1} \; k, \; w' \;
  \equiv_{q_1} \; k', \text{ and } k \; \equiv_{q_2} \; k'\\
      w \; \equiv_{q_1 \cdot q_2} \; w'  &\iff& w = r_1
      \cdot r_2, \; w' = {r'}_1 \cdot {r'}_2 \text{ with } r_1 \; \equiv_{q_1}
      \; {r'}_1, \; r_2 \; \equiv{q_n} \; {r'}_2\\
      w \; \equiv_{q' \sep q} \; w' &\iff& w \; \equiv_{q_1} \; w'
      \text{ or } \; w \; \equiv_{q_2} \; w'\\
      w \; \equiv_{q*} \; w' &\iff& w = r_1 \cdot \ldots \cdot r_n, \; w'
      = {r'}_1 \cdot \ldots \cdot {r'}_n \text{ and } r_i \equiv_{q} \; {r'}_i
      \\
      w \; \equiv_{\perm{q_1, \ldots, q_n}{q}} \; w' &\iff& w = r_{\sigma(1)}
      \cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}, \;
    w' = {r'}_{\theta(1)} \cdot s'_1 \cdot \ldots \cdot s'_{n-1}
    \cdot {r'}_{\theta(n)} \\
    & & \text{ for some } \sigma, \theta \in S_n, \text{ with } r_i \;
    \equiv_{q_i} \; r'_i \text{ and } s_k \; \equiv_{q} \; s'_{k}
    \end{array}
    \]
  \caption{QRE Equivalence Relations}
  \label{fig:relations}
\end{figure}
\begin{figure}[t]
  \begin{center}
\[
    \begin{array}{l@\quad l @\quad l} 
      \canonizer(R) &=& id_{\mathcal{L}(R)} \\
      \canonizer(R \mapsto s)(w) &=& s\\
      \canonizer(\squash{R}{R'}{f})(w) &=& 
\begin{cases}
f(w) & \text{if } w \in \mathcal{L}(R)\\
w & \text{otherwise}
\end{cases}\\
      \canonizer(\normalize{R}{R'}{f}) &=& f\\
      \canonizer(q' \circ q) &=& \canonizer(q') \circ \canonizer(q)\\
      \canonizer(q' \cdot q) &=& \canonizer(q') \cdot \canonizer(q)\\
      \canonizer(q_1 \sep q_2)(w) &=& 
\begin{cases}
\canonizer(q_1)(w) & \text{if } w \in \mathcal{L}(W(q_1))\\
\canonizer(q_2)(w) & \text{if } w \in \mathcal{L}(W(q_2))\\
\end{cases}\\
      \canonizer(q^*) &=& \canonizer(q)^* \\
    \end{array}
    \]
    \end{center}
    $\canonizer(\perm{q_1, \ldots, q_n}{q})(r_{\sigma(1)}
\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}) \newline
= (\canonizer(q_1)(r_1)) \cdot (\canonizer(q)(s_1)) \cdot \ldots \cdot
(\canonizer(q_n)(r_n))$
  \caption{QRE canonizers}
  \label{fig:canonizers}
\end{figure}
\begin{figure}[t]
\centering
\begin{prooftree}
\AxiomC{$R$ is strongly unambiguous}
\UnaryInfC{$\mathit{id}(R)$ is well formed}
\end{prooftree}

@@

\begin{prooftree}
\AxiomC{$q$ is well formed}
\AxiomC{$W(q)^{*!}$}
\AxiomC{$K(q)^{*!}$}
\TrinaryInfC{$q^*$ is well formed}
\end{prooftree}
  \caption{QRE Inference Rules\bcp{Maybe some or all of these can be
      suppressed in the ``short version'' of the paper}\bcp{No, cancel
      that.  These are important.  And we need to comment about the decidability
    issues involving normalize.  (And a note that we don't really use
    normalize, but rather introduce more specialized instances.)}}
  \label{fig:qrerules}
\end{figure}
The theorem below confirms that the semantics of QREs are consistent with their
intended behaviour:
\begin{theorem}
If $q$ is a well formed QRE, then
\begin{enumerate}
  \item $W(q)$ and $K(q)$ are well defined regular expressions,
  \item  $\eqrel{q}$ is an equivalence relation on $\mathcal{L}(W(q))$,
  \item  $\mathcal{L}(K(q))$ forms a complete set of representatives for
  $\eqrel{q}$,
  \item $\canonizer(q):\mathcal{L}(W(q)) \longrightarrow \mathcal{L}(K(q))$ is a
  well-defined function, and
  \item  given any $w \in \mathcal{L}(W(q))$, $\canonizer(q)(w)$ is the unique
  $k$ in $\mathcal{L}(K(q))$ such that $k$ is equivalent to $w$ modulo
  $\eqrel{q}$.
  \end{enumerate}
\end{theorem}

\section{QRE Lenses}
\label{QRE-lenses} 
As we have seen, QREs express a broad class of equivalence relations
directly on regular languages.  QREs are therefore a good \textit{specification}
language for quotient lenses. In this section of the paper, we introduce
\textit{QRE Lenses}, a class of quotient lenses specified using QREs.

Our approach in defining QRE lenses is as follows. Let $c, c'$ be QREs. Given a
bijection $\ell : K(c) \Leftrightarrow K(c')$, we may define a quotient lens $q
: W(c)/\eqrel{c} \Leftrightarrow W(c')/\eqrel{c'}$ by
\begin{equation}\label{normalform}
q.get = \ell \circ \canonizer(c) \text{ and } q.put = \ell^{-1} \circ
\canonizer(c'),
\end{equation}

@@

One potential pitfall with QRE lenses is that the combinators which
we define on them may not preserve this normal form. For example, the
composition $q' \circ q$ of QRE lenses $q$ and $q'$ may not be equivalent to
any quotient lens which has the canonical normal form of a bijective lens with
canonizers a the edges. Consequently, in order to define quotient lenses, we
need to choose a set $S$ of bijections that has the right properties in the
sense that QRE lenses built from the bijections in $S$ are closed under the
action of QRE lens combinators. Our choice is to use the class of
\textit{bijective lenses}, which are the bijections that can be expressed in
Boomerang. In the next subsection we briefly introduce bijective lenses and
discuss the properties of bijective lenses which we will use to prove that QRE
lenses have the normal form of bijective lens with canonizers at the edges.

@@

\begin{align*}
\llbracket R \rrbracket &= \{(r, r) \sep r \in \mathcal{L}(R)\}\\
\llbracket swap(\ell, \ell') \rrbracket &= \{(s \cdot t, t' \cdot s') \sep
(s, s') \in \llbracket \ell \rrbracket \text{ and } (t, t') \in \llbracket
\ell' \rrbracket\}\\
\llbracket \ell \cdot \ell' \rrbracket &= \{(s \cdot t, s' \cdot t) \sep
(s, s') \in \llbracket \ell \rrbracket \text{ and } (t, t') \in \llbracket
\ell' \rrbracket\}\\
\llbracket \ell \sep \ell' \rrbracket &= \{(s \cdot t) \sep
(s, t) \in \llbracket \ell \rrbracket \text{ or } (s, t) \in \llbracket
\ell' \rrbracket\}\\
\llbracket \ell^* \rrbracket &= \{(s_1 \cdot \ldots \cdot s_n, t_1 \cdot \ldots
\cdot t_n) \sep (s_i, t_i) \in \llbracket \ell \rrbracket \text{ for } 1
\leq i \leq n\}
\end{align*}

@@

Each bijective lens $\ell$ has a type $\ell : R \Leftrightarrow S$ where $R$ and
$S$ are regular expressions. If $\ell : R \Leftrightarrow S$, then the source
type of $\ell$ is $\mathcal{L}(R)$ and the target type of $\ell$ is
$\mathcal{L}(S)$. Interestingly, with the bijective lens type system, a lens
$\ell : R \Leftrightarrow S$ can also be considered to be of type $\ell : R'
\Leftrightarrow S'$ provided that $R$ (resp. $S$) can be proven to be
equivalent to $R'$ (resp. $S$) from the star-semiring axioms:

@@

The more interesting combinators are the $\mathit{lquot}$ and $\mathit{rquot}$
combinators. The $\mathit{lquot}(c, q)$ combinator takes a quotient lens $q$ and
quotients the source data using $c$, assuming that the source data forms a
complete set of representatives for the equivalence relation $\eqrel{c}$. The
$\mathit{lquot}(q, c)$ combinator does the same but on the view data.

@@

\begin{lstlisting}
let bib_to_end_q : (lens in bibtex <=> endnote) = 
rquot (lquot (bibtex, bib_to_end), endnote)
 \end{lstlisting}

@@

\subsection{Semantics of QRE Lenses}
Each QRE lens $q$ has a type $q : c \Leftrightarrow c'$ where $c, c'$ are QREs.
If $q : c \Leftrightarrow c'$, then the source format is described by $W(c)$ and
the canonical set of representative for the source data is described by $K(c)$.
Similarly, the target format is described by $W(c')$ and the canonical set of
representatives for the target data is described by $K(c')$. The underlying lens
of $q$ is a bijective lens $\ell : K(c) \Leftrightarrow K(c')$.

@@

 \begin{figure}[t]
  \centering
  \begin{prooftree}
\AxiomC{$\ell : R \Leftrightarrow S$}
\UnaryInfC{$\mathit{lift}(\ell): id(R) \Leftrightarrow id(S)$}
\end{prooftree}
  \begin{align*}
  \llbracket q \rrbracket.get &=  \llbracket \ell \rrbracket, \text{ and }\\
  \llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1}
  \end{align*}

@@

\begin{theorem}\label{normal form}
If there is a derivation $q : c \Leftrightarrow c'$, then there exists a
bijective lens $\ell : K(c) \Leftrightarrow K(c')$ such that
\begin{align*}
\llbracket q \rrbracket.get &= \llbracket \ell \rrbracket\circ \canonizer(c)\\
\llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
\canonizer(c')
\end{align*}
\end{theorem}
We elide a full proof of this statement, since the proof is attained via a
straightforward induction on the derivation $q : c \Leftrightarrow c$. Since
bijective lenses are closed under strongly unambiguous regular operators, the
tricky case is composition. The composition case is more involved since we need
to show that we can get rid of the canonizers in the middle:

@@

The other cases of the proof are similar in that they follow from a
straightforward application of the induction hypothesis followed by unrolling
the equations that give the denotation for QRE lenses.
\iffalse
\begin{proof}
Assume that $q : c \Leftrightarrow c'$. We proceed by induction over the
derivation $q : c \Leftrightarrow c'$.
\begin{enumerate}
  \item
  $\mathit{lift}(\ell): R/\mathit{id}(R) \Leftrightarrow S/\mathit{id}(S)$ where
  $\ell :
  R \Leftrightarrow S$. Then
  \begin{align*}
  \llbracket \mathit{lift}(\ell) \rrbracket.get &=  \llbracket \ell \rrbracket
  = \llbracket \ell \rrbracket \circ id_{\mathcal{L}(R)} =
  \llbracket \ell \rrbracket \circ \canonizer(\mathit{id}(R)), \text{ and }\\
  \llbracket \mathit{lift}(\ell) \rrbracket.put &= \llbracket \ell
  \rrbracket^{-1} = \llbracket \ell \rrbracket^{-1} \circ id_{\mathcal{L}(S)} =
  \llbracket \ell \rrbracket^{-1} \circ \canonizer(id(S))
  \end{align*}
  \item
  $\mathit{lquot}(c, q'): c' \circ c \Leftrightarrow c''$ where $q' : c' 
  \Leftrightarrow c''$, $c$ is well formed and $K(c) = W(c')$. Then
\begin{align*}
  \llbracket q \rrbracket.get  &= \llbracket q'
  \rrbracket.get \circ \canonizer(c)\\
  \llbracket q \rrbracket.put &= \llbracket q' \rrbracket.put
  \end{align*}
  By the induction hypothesis, there exists a bijective lens $\ell :
  K(c') \Leftrightarrow K(c'')$ such that 
  \begin{align*}
\llbracket q' \rrbracket.get &= \llbracket \ell \rrbracket \circ \canonizer(c')\\
\llbracket q' \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
Canonize(c'')
\end{align*}
Consequently
\begin{align*}
  \llbracket q \rrbracket.get  &= (\llbracket \ell \rrbracket \circ
  \canonizer(c')) \circ \canonizer(c) = \llbracket \ell \rrbracket \circ
  (\canonizer(c' \circ c))\\
  \llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
  Canonize(c'')
  \end{align*}

>>>>

\saz{Is this where we should explain the rules of  Figure~\ref{fig:qrerules}?}

@@

\begin{figure}[t]
  \centering
\saz{Fix the projection case}
\[
    \begin{array}{l@{\quad}l@{\quad}l} 
      w \; \equiv_R \; w' &\iff& w = w' \\
      w \; \equiv_{R \mapsto s} \; w' \text{ for all }w, w'\\
      w \; \equiv_{\squash{R}{R'}{f}} \; w' &\iff& f(w) = w'
      \text{ or } w = w' \\
      w \; \equiv_{\normalize{R}{R'}{f}} \; w' &\iff&
      f(w)=f(w') \text{ or }w = w'\\
      w \; \equiv_{q_2 \circ q_1} \; w' &\iff& \exists k, k' \in
  \mathcal{L}(K(q_2)) \text{ such that } w \; \equiv_{q_1} \; k, \; w' \;
  \equiv_{q_1} \; k', \text{ and } k \; \equiv_{q_2} \; k'\\
      w \; \equiv_{q_1 \cdot q_2} \; w'  &\iff& w = r_1
      \cdot r_2, \; w' = {r'}_1 \cdot {r'}_2 \text{ with } r_1 \; \equiv_{q_1}
      \; {r'}_1, \; r_2 \; \equiv{q_n} \; {r'}_2\\
      w \; \equiv_{q' \sep q} \; w' &\iff& w \; \equiv_{q_1} \; w'
      \text{ or } \; w \; \equiv_{q_2} \; w'\\
      w \; \equiv_{q*} \; w' &\iff& w = r_1 \cdot \ldots \cdot r_n, \; w'
      = {r'}_1 \cdot \ldots \cdot {r'}_n \text{ and } r_i \equiv_{q} \; {r'}_i
      \\
      w \; \equiv_{\perm{q_1, \ldots, q_n}{q}} \; w' &\iff& w = r_{\sigma(1)}
      \cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}, \;
    w' = {r'}_{\theta(1)} \cdot s'_1 \cdot \ldots \cdot s'_{n-1}
    \cdot {r'}_{\theta(n)} \\
    & & \text{ for some } \sigma, \theta \in S_n, \text{ with } r_i \;
    \equiv_{q_i} \; r'_i \text{ and } s_k \; \equiv_{q} \; s'_{k}
    \end{array}
    \]
  \caption{QRE Equivalence Relations}
  \label{fig:relations}
\end{figure}
\begin{figure}[t]
  \begin{center}
\[
    \begin{array}{l@\quad l @\quad l} 
      \canonizer(R) &=& id_{\mathcal{L}(R)} \\
      \canonizer(R \mapsto s)(w) &=& s\\
      \canonizer(\squash{R}{R'}{f})(w) &=& 
\begin{cases}
f(w) & \text{if } w \in \mathcal{L}(R)\\
w & \text{otherwise}
\end{cases}\\
      \canonizer(\normalize{R}{R'}{f}) &=& f\\
      \canonizer(q' \circ q) &=& \canonizer(q') \circ \canonizer(q)\\
      \canonizer(q' \cdot q) &=& \canonizer(q') \cdot \canonizer(q)\\
      \canonizer(q_1 \sep q_2)(w) &=& 
\begin{cases}
\canonizer(q_1)(w) & \text{if } w \in \mathcal{L}(W(q_1))\\
\canonizer(q_2)(w) & \text{if } w \in \mathcal{L}(W(q_2))\\
\end{cases}\\
      \canonizer(q^*) &=& \canonizer(q)^* \\
    \end{array}
    \]
    \end{center}
    $\canonizer(\perm{q_1, \ldots, q_n}{q})(r_{\sigma(1)}
\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}) \newline
= (\canonizer(q_1)(r_1)) \cdot (\canonizer(q)(s_1)) \cdot \ldots \cdot
(\canonizer(q_n)(r_n))$
  \caption{QRE canonizers}
  \label{fig:canonizers}
\end{figure}

\begin{figure}[p!]
\centering
\begin{prooftree}
\AxiomC{$R$ is strongly unambiguous}
\UnaryInfC{$\mathit{id}(R)$ is well formed}
\end{prooftree}

@@

\begin{prooftree}
\AxiomC{$q$ is well formed}
\AxiomC{$W(q)^{*!}$}
\AxiomC{$K(q)^{*!}$}
\TrinaryInfC{$q^*$ is well formed}
\end{prooftree}
  \caption{Well-formed QREs \bcp{Maybe some or all of these can be
      suppressed in the ``short version'' of the paper}\bcp{No, cancel
      that.  These are important.  And we need to comment about the decidability
    issues involving normalize.  (And a note that we don't really use
    normalize, but rather introduce more specialized instances.)}
  \saz{maybe shorten the judgment from ``is well formed'' to ``wf''?}
}
  \label{fig:qrerules}
\end{figure}

\saz{We need to explain the rules in Figure~\ref{fig:qrerules}---otherwise they
  aren't mentioned anywhere in the text!}
The theorem below confirms that the semantics of QREs are consistent with their
intended behaviour:
\begin{theorem}
If $q$ is a well formed QRE, then
\begin{enumerate}
  \item $W(q)$ and $K(q)$ are well defined regular expressions,
  \item  $\eqrel{q}$ is an equivalence relation on $\mathcal{L}(W(q))$,
  \item  $\mathcal{L}(K(q))$ forms a complete set of representatives for
  $\eqrel{q}$,
  \item $\canonizer(q):\mathcal{L}(W(q)) \longrightarrow \mathcal{L}(K(q))$ is a
  well-defined function, and
  \item  given any $w \in \mathcal{L}(W(q))$, $\canonizer(q)(w)$ is the unique
  $k$ in $\mathcal{L}(K(q))$ such that $k$ is equivalent to $w$ modulo
  $\eqrel{q}$.
  \end{enumerate}
\end{theorem}

\section{QRE Lenses}
\label{QRE-lenses}

\saz{Throughout this section: do a which hunt to look for ``which''es that
  should be ``that''s.}

As we have seen, QREs express a broad class of equivalence relations
directly on regular languages.  QREs are therefore a good \textit{specification}
language for quotient lenses. In this section of the paper, we introduce
\textit{QRE Lenses}, a class of quotient lenses specified using QREs.

\saz{We need to be careful with our metavariable conventions.  Up until now $q$
  has meant ``a QRE'', but this sections says that $q$ and $q'$ are
  \textit{lenses} and switches to using $c$ and $c'$ for QREs. We need to
  pick a metavariable that ranges over QRE lenses and use it consistently and
  switch these $c$'s to $q$'s.
}

Our approach in defining QRE lenses is as follows. Let $c$ and $c'$ be QREs. Given a
bijection $\ell : K(c) \Leftrightarrow K(c')$, we may define a quotient lens $q
: W(c)/\eqrel{c} \Leftrightarrow W(c')/\eqrel{c'}$ by
\begin{equation}\label{normalform}
q.get = \ell \circ \canonizer(c) \text{ and } q.put = \ell^{-1} \circ
\canonizer(c'),
\end{equation}

@@

One potential pitfall with QRE lenses is that the combinators that
we define on them may not preserve this normal form. For example, the
composition $q' \circ q$ of QRE lenses $q$ and $q'$ may not be equivalent to
any quotient lens which has the canonical normal form of a bijective lens with
canonizers a the edges. Consequently, in order to define quotient lenses, we
need to choose a set $S$ of bijections that has the right properties in the
sense that QRE lenses built from the bijections in $S$ are closed under the
action of QRE lens combinators. Our choice is to use the class of
\textit{bijective lenses}, which are the bijections that can be expressed in
Boomerang. In the next subsection we briefly introduce bijective lenses and
discuss the properties of bijective lenses which we will use to prove that QRE
lenses have the normal form of bijective lens with canonizers at the edges.

@@

\saz{include the denotations of const and $\circ$}
\begin{align*}
\llbracket R \rrbracket &= \{(r, r) \sep r \in \mathcal{L}(R)\}\\
\llbracket swap(\ell, \ell') \rrbracket &= \{(s \cdot t, t' \cdot s') \sep
(s, s') \in \llbracket \ell \rrbracket \text{ and } (t, t') \in \llbracket
\ell' \rrbracket\}\\
\llbracket \ell \cdot \ell' \rrbracket &= \{(s \cdot t, s' \cdot t) \sep
(s, s') \in \llbracket \ell \rrbracket \text{ and } (t, t') \in \llbracket
\ell' \rrbracket\}\\
\llbracket \ell \sep \ell' \rrbracket &= \{(s \cdot t) \sep
(s, t) \in \llbracket \ell \rrbracket \text{ or } (s, t) \in \llbracket
\ell' \rrbracket\}\\
\llbracket \ell^* \rrbracket &= \{(s_1 \cdot \ldots \cdot s_n, t_1 \cdot \ldots
\cdot t_n) \sep (s_i, t_i) \in \llbracket \ell \rrbracket \text{ for } 1
\leq i \leq n\}
\end{align*}

@@

Each bijective lens $\ell$ has a type $\ell : R \Leftrightarrow S$ where $R$ and
$S$ are regular expressions. If $\ell : R \Leftrightarrow S$, then the source
language of $\ell$ is $\mathcal{L}(R)$ and the target language of $\ell$ is
$\mathcal{L}(S)$. Interestingly, with the bijective lens type system, a lens
$\ell : R \Leftrightarrow S$ can also be considered to be of type $\ell : R'
\Leftrightarrow S'$ provided that $R$ (resp. $S$) can be proven to be
equivalent to $R'$ (resp. $S$) from the star-semiring axioms:

@@

\saz{Watch metavariables}
The more interesting combinators are the $\mathit{lquot}$ and $\mathit{rquot}$
combinators. The $\mathit{lquot}(c, q)$ combinator takes a quotient lens $q$ and
quotients the source data using $c$, assuming that the source data forms a
complete set of representatives for the equivalence relation $\eqrel{c}$. The
$\mathit{lquot}(q, c)$ combinator does the same but on the view data.

@@

\begin{lstlisting}
let bib_to_end_q : (lens in bibtex $\Leftrightarrow$ endnote) = 
rquot (lquot (bibtex, bib_to_end), endnote)
 \end{lstlisting}

@@

\subsection{Semantics of QRE Lenses}

\saz{metavariables\ldots}
Each QRE lens $q$ has a type $q : c \Leftrightarrow c'$ where $c, c'$ are QREs.
If $q : c \Leftrightarrow c'$, then the source format is described by $W(c)$ and
the canonical set of representative for the source data is described by $K(c)$.
Similarly, the target format is described by $W(c')$ and the canonical set of
representatives for the target data is described by $K(c')$. The underlying lens
of $q$ is a bijective lens $\ell : K(c) \Leftrightarrow K(c')$.

@@

 \begin{figure}[ht]
  \centering
  \begin{prooftree}
\AxiomC{$\ell : R \Leftrightarrow S$}
\UnaryInfC{$\mathit{lift}(\ell): id(R) \Leftrightarrow id(S)$}
\end{prooftree}
  \begin{align*}
  \llbracket q \rrbracket.get &=  \llbracket \ell \rrbracket, \text{ and }\\
  \llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1}
  \end{align*}

@@

\begin{theorem}\label{normal form}
If there is a derivation $q : c \Leftrightarrow c'$, then there exists a
bijective lens $\ell : K(c) \Leftrightarrow K(c')$ such that
\begin{align*}
\llbracket q \rrbracket.get &= \llbracket \ell \rrbracket\circ \canonizer(c)\\
\llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
\canonizer(c')
\end{align*}
\end{theorem}
\begin{proof} 
We elide a full proof of this statement, since the proof is attained via a
straightforward induction on the derivation $q : c \Leftrightarrow c$. \saz{(The
  full details can be found in the appendix?)}Since
bijective lenses are closed under strongly unambiguous regular operators, the
tricky case is composition. The composition case is more involved since we need
to show that we can get rid of the canonizers in the middle:

@@

The other cases of the proof are similar in that they follow from a
straightforward application of the induction hypothesis followed by unrolling
the equations that give the denotation for QRE lenses.\qed
\end{proof}
\iffalse
\begin{proof}
Assume that $q : c \Leftrightarrow c'$. We proceed by induction over the
derivation $q : c \Leftrightarrow c'$.
\begin{enumerate}
  \item
  $\mathit{lift}(\ell): R/\mathit{id}(R) \Leftrightarrow S/\mathit{id}(S)$ where
  $\ell :
  R \Leftrightarrow S$. Then
  \begin{align*}
  \llbracket \mathit{lift}(\ell) \rrbracket.get &=  \llbracket \ell \rrbracket
  = \llbracket \ell \rrbracket \circ id_{\mathcal{L}(R)} =
  \llbracket \ell \rrbracket \circ \canonizer(\mathit{id}(R)), \text{ and }\\
  \llbracket \mathit{lift}(\ell) \rrbracket.put &= \llbracket \ell
  \rrbracket^{-1} = \llbracket \ell \rrbracket^{-1} \circ id_{\mathcal{L}(S)} =
  \llbracket \ell \rrbracket^{-1} \circ \canonizer(id(S))
  \end{align*}
  \item
  $\mathit{lquot}(c, q'): c' \circ c \Leftrightarrow c''$ where $q' : c' 
  \Leftrightarrow c''$, $c$ is well formed and $K(c) = W(c')$. Then
\begin{align*}
  \llbracket q \rrbracket.get  &= \llbracket q'
  \rrbracket.get \circ \canonizer(c)\\
  \llbracket q \rrbracket.put &= \llbracket q' \rrbracket.put
  \end{align*}
  By the induction hypothesis, there exists a bijective lens $\ell :
  K(c') \Leftrightarrow K(c'')$ such that 
  \begin{align*}
\llbracket q' \rrbracket.get &= \llbracket \ell \rrbracket \circ \canonizer(c')\\
\llbracket q' \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
Canonize(c'')
\end{align*}
Consequently
\begin{align*}
  \llbracket q \rrbracket.get  &= (\llbracket \ell \rrbracket \circ
  \canonizer(c')) \circ \canonizer(c) = \llbracket \ell \rrbracket \circ
  (\canonizer(c' \circ c))\\
  \llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
  Canonize(c'')
  \end{align*}
